---
title: "Byte pair encoding tokenization for geographical place names"
author: "Mike Kaminski"
date: "2023-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE,
                      message = FALSE, dpi = 180)
```

```{r}
library(tidyverse)

us_place_names <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-27/us_place_names.csv')

glimpse(us_place_names)
```

```{r}
place_counts <- us_place_names |>
  count(feature_name, sort = TRUE) |>
  filter(n >1)

```

```{r}
place_counts |>
  ggplot(aes(n)) +
  geom_histogram(bins = 15) +
  scale_x_log10()
  
```

## Build a model
```{r}
library(tidymodels)

set.seed(54)
place_split <- initial_split(place_counts, strata = n)
place_train <- training(place_split)
place_test <- testing(place_split)
```

Builds a feature engineering recipe
```{r}
library(textrecipes)
library(tokenizers.bpe)

place_rec <-
  recipe(n ~ feature_name, data = place_train) |>
  step_tokenize_bpe(feature_name, vocabulary_size = 200) |> # byte pair encoding - we need to tune this
  step_tokenfilter(feature_name, max_tokens = 100) |>
  step_tf(feature_name) #term frequency

```

Most large language models use byte pair encoding tokenization

```{r}
prep(place_rec) |> bake(new_data = NULL) |> glimpse()

```
term frequency of each feature name

Byte pair encoding iteratively merges frequently occuring words/sub words to decide what the vocabulary.  It merges sub word information to find the most important subwords.

```{r}
library(poissonreg)
poisson_wf <- workflow(place_rec, poisson_reg())

poisson_fit <- (fit(poisson_wf,place_train))
```



## Understand the results
```{r}
library(stringr)
tidy(poisson_fit) |>
  filter(term != "(Intercept)") |>
  mutate(term = str_remove(term, "tf_feature_name_")) |>
  slice_max(abs(estimate), n = 20) |>
  arrange(-estimate)

```

```{r}
place_train |>
  filter(str_detect(feature_name, "Estates|wood")) |>
  mutate(feature_name = case_when(
    str_detect(feature_name,"wood") ~ "wood",
    str_detect(feature_name, "Estate") ~ "estate"
  ))|>
  ggplot(aes(n, fill = feature_name)) +
  geom_histogram(alpha = 0.8, position = "dodge", bins = 15) +
  scale_x_log10()

```
More unique estate names, more common wood names















