---
title: "K-means clustering with tidy data principles"
author: "Mike Kaminski"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective: summarize clustering characteristics and estimate the number of clusters

# Introduction:
- use tidy, augment, and glance

```{r}
# random data with three clusters.  Each will come from a multivariate gaussian distribution with differnet means for each cluster

library(tidymodels)

set.seed(525)

centers <- tibble(
  cluster = factor(1:3),
  num_points = c(100, 150, 50),
  x1 = c(5 ,0, -3),
  x2 = c(-1, 1, -2)
)

labelled_points <-
  centers %>%
  mutate(
    x1 = map2(num_points, x1, rnorm),
    x2 = map2(num_points, x2, rnorm)
  ) %>%
  select(-num_points) %>%
  unnest(cols = c(x1, x2))

ggplot(labelled_points, aes(x1, x2, color = cluster)) +
  geom_point(alpha = 0.3)

```
# How does K-Means work?
- specify the number of clusters
- centroids are randomly placed
- each observation gets assigned to the nearest centroid based on Euclidean distance.
- then the centroid is recalculated based on the clusters
- then each point is reassigned to the nearest cluster
- then the centroid is recalculated based on the clusters
- then each point is reassigned to the nearest cluster
-.....
- iterative process until no changes.

```{r}
points <-
  labelled_points %>%
  select(-cluster)

kclust <- kmeans(points, center = 3)
kclust

summary(kclust)

```
This gives us a list of vectors
- cluster 300 values, which is the data about each point.
- centers, withinss, and size contrain info about each cluster
-- centers includes cluster means at x1 and x2 for each cluster
- totss, tot.withinss, betweenss, and iter contain info about the full clustering

```{r}
# this gives us each point of data
augment(kclust, points)
```

```{r}
#thi gives us a per-cluster level
tidy(kclust)

```

```{r}
# this gives us a single row summary
glance(kclust)

```
# Exploratory Clustering
We want to combine this power with dplyr
We can explore different choices of k, from 1 to 9.
- We cluster 9 times, with each having a different value of k.  Then we create columns containing the tidied, glanced, and augmented data
```{r}
kclusts <-
  tibble(k = 1:9) %>%
  mutate(kclust = map(k, ~kmeans(points, .x)),
         tidied = map(kclust, tidy),
         glanced = map(kclust, glance),
         augmented = map(kclust, augment, points)
         )
kclusts

```

Now we can turn into 3 seprate datasets for tidy, augment, and glance
```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))


```

And we plot
```{r}
p1 <-
  ggplot(assignments, aes(x = x1, y = x2)) +
  geom_point(aes(color = .cluster), alpha = 0.8) +
  facet_wrap(~k)
p1


```

It's pretty clear that 3 is the optimal number of clusters.  We can add an X to the center of each cluster
```{r}
p2 <- p1 + geom_point(data = clusters, size = 10, shape = "x")
p2

```

The data from glance fills a bit of a different purpose.  We want to look at total within sum of squares.
- This represents the variance within the clusters.  We want to look for the elbow bend.
```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()

```
```{r}
sessionInfo()


```
attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] yardstick_1.1.0    workflowsets_1.0.0 workflows_1.1.2    tune_1.0.1         tidyr_1.2.1       
 [6] tibble_3.1.8       rsample_1.1.0      recipes_1.0.3      purrr_0.3.5        parsnip_1.0.3     
[11] modeldata_1.0.1    infer_1.0.4        ggplot2_3.4.1      dplyr_1.0.10       dials_1.1.0       
[16] scales_1.2.1       broom_1.0.1        tidymodels_1.0.0  

loaded via a namespace (and not attached):
 [1] pkgload_1.3.2       viridisLite_0.4.1   splines_4.2.2       foreach_1.5.2      
 [5] prodlim_2019.11.13  assertthat_0.2.1    GPfit_1.0-8         yaml_2.3.6         
 [9] globals_0.16.2      ipred_0.9-13        pillar_1.8.1        backports_1.4.1    
[13] lattice_0.20-45     glue_1.6.2          digest_0.6.30       hardhat_1.2.0      
[17] colorspace_2.0-3    htmltools_0.5.5     Matrix_1.5-1        timeDate_4021.106  
[21] pkgconfig_2.0.3     lhs_1.1.5           DiceDesign_1.9      listenv_0.8.0      
[25] gower_1.0.0         lava_1.7.0          timechange_0.1.1    farver_2.1.1       
[29] generics_0.1.3      ellipsis_0.3.2      withr_2.5.0         furrr_0.3.1        
[33] nnet_7.3-18         cli_3.6.0           survival_3.4-0      magrittr_2.0.3     
[37] evaluate_0.18       future_1.29.0       fansi_1.0.3         parallelly_1.32.1  
[41] MASS_7.3-58.1       class_7.3-20        tools_4.2.2         lifecycle_1.0.3    
[45] munsell_0.5.0       compiler_4.2.2      rlang_1.0.6         grid_4.2.2         
[49] iterators_1.0.14    rstudioapi_0.14     labeling_0.4.2      rmarkdown_2.18     
[53] gtable_0.3.1        codetools_0.2-18    DBI_1.1.3           R6_2.5.1           
[57] lubridate_1.9.0     knitr_1.41          fastmap_1.1.0       future.apply_1.10.0
[61] utf8_1.2.2          parallel_4.2.2      Rcpp_1.0.9          vctrs_0.5.1        
[65] rpart_4.1.19        tidyselect_1.2.0    xfun_0.35   
