---
title: "Correlation and regression fundamentals with tidy data principles"
author: "Mike Kaminski"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective: Analyze the results of correlation tests and simple regression models for many data sets at once

# Intro:
The broom package in tidymodels is useful for a single analysis, however it's not useful for combining multiple analyses.
- subgroups of data, analyses using diff models, bootstraps, permutations, etc.


# Correlation Analysis:
Note: we coerce into a tibble so that we get a nicer print output
```{r}
library(tidymodels)

data(Orange)
Orange <- as_tibble(Orange)
Orange

```
Tree, age, and circumference are the variables.  Tree describes 5 types of trees

```{r}
cor(Orange$age, Orange$circumference)

library(ggplot2)

ggplot(Orange, aes(age, circumference, color = Tree)) +
  geom_line()

```

We can test the correltaion within each tree
```{r}
Orange %>%
  group_by(Tree) %>%
  summarize(correlation = cor(age, circumference))

```

Now let's say we want to perform a hypothesis test instead of simply estimating the correlation
```{r}
ct <- cor.test(Orange$age, Orange$circumference)
ct

tidy(ct)
```
Usually we want to perform multiple tests or fit multiple models on different parts of the data.  Here we can use a nest-map-unnest workflow.

We want to perform correlation tests for each different tree
```{r}
# start by nesting
nested <-
  Orange %>%
  nest(data = c(age, circumference))

# then we do a correlation test for each tibble using purrr::map()
nested %>%
  mutate(test = map(data, ~ cor.test(.x$age, .x$circumference)))

# tidy each of the objects
nested %>%
  mutate(test = map(data, ~ cor.test(.x$age, .x$circumference)), # S# list-col
         tidied = map(test, tidy)
         )

# then we want to unnest the tidied dataframe so that the results are a flat tibble.

Orange %>% 
  nest(data = c(age, circumference)) %>% 
  mutate(
    test = map(data, ~ cor.test(.x$age, .x$circumference)), # S3 list-col
    tidied = map(test, tidy)
  ) %>% 
  unnest(cols = tidied) %>% 
  select(-data, -test)
```

# Regression Models
This type of workflow is more useful when applied to regressions

```{r}

options(scipen = 10, digits =4)
lm_fit <- lm(age ~ circumference, data = Orange)
summary(lm_fit)
tidy(lm_fit)
```

We can do multiple regressions at once using the same workflow
```{r}
Orange %>%
  nest(data = c(-Tree)) %>%
  mutate(
    fit = map(data, ~ lm(age ~ circumference, data = .x)),
    tidied = map(fit, tidy)
  ) %>%
  unnest(tidied) %>%
  select(-data, -fit)
  


```

We can also use multiple predictors in the regressions.  We nest the data into automatic vs manual and then do regression
```{r}
data(mtcars)
mtcars <- as_tibble(mtcars)
mtcars

mtcars %>%
  nest(data = c(-am)) %>%
  mutate(
    fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),  # S3 list-col
    tidied = map(fit, tidy)
  ) %>%
  unnest(tidied) %>%
  select(-data, -fit)

```


We can have the tidy(), as well as the aug,ent() and glance()
```{r}
regressions <- 
  mtcars %>%
    nest(data = c(-am)) %>%
    mutate(
      fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),  # S3 list-col
      tidied = map(fit, tidy),
      glanced = map(fit, glance),
      augmented = map(fit, augment)
    )

regressions %>%
  select(tidied) %>%
  unnest(tidied)

regressions %>%
  select(glanced) %>%
  unnest(glanced)


regressions %>%
  select(augmented) %>%
  unnest(augmented)

```

When we combine the estimates and p-values across all groups into the same tidy dataframe - instead of a list of output model objects - class analyses and visualizations become more straighforward
- we can ort p-values or estimate to find the most significant terms across all tests
- we can make p-value histograms
- we can make volcano plots comparing p-values to effect size estimates.

We can easily filter, facet, or distinguish based on the 'term' column

```{r}
sessionInfo()


```
attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] yardstick_1.1.0    workflowsets_1.0.0 workflows_1.1.2    tune_1.0.1         tidyr_1.2.1        tibble_3.1.8       rsample_1.1.0     
 [8] recipes_1.0.3      purrr_0.3.5        parsnip_1.0.3      modeldata_1.0.1    infer_1.0.4        ggplot2_3.4.1      dplyr_1.0.10      
[15] dials_1.1.0        scales_1.2.1       broom_1.0.1        tidymodels_1.0.0  

loaded via a namespace (and not attached):
 [1] pkgload_1.3.2       viridisLite_0.4.1   splines_4.2.2       foreach_1.5.2       prodlim_2019.11.13  assertthat_0.2.1    GPfit_1.0-8        
 [8] yaml_2.3.6          globals_0.16.2      ipred_0.9-13        pillar_1.8.1        backports_1.4.1     lattice_0.20-45     glue_1.6.2         
[15] digest_0.6.30       hardhat_1.2.0       colorspace_2.0-3    htmltools_0.5.5     Matrix_1.5-1        timeDate_4021.106   pkgconfig_2.0.3    
[22] lhs_1.1.5           DiceDesign_1.9      listenv_0.8.0       gower_1.0.0         lava_1.7.0          timechange_0.1.1    farver_2.1.1       
[29] generics_0.1.3      ellipsis_0.3.2      withr_2.5.0         furrr_0.3.1         nnet_7.3-18         cli_3.6.0           survival_3.4-0     
[36] magrittr_2.0.3      evaluate_0.18       future_1.29.0       fansi_1.0.3         parallelly_1.32.1   MASS_7.3-58.1       class_7.3-20       
[43] tools_4.2.2         lifecycle_1.0.3     munsell_0.5.0       compiler_4.2.2      rlang_1.0.6         grid_4.2.2          iterators_1.0.14   
[50] rstudioapi_0.14     labeling_0.4.2      rmarkdown_2.18      gtable_0.3.1        codetools_0.2-18    DBI_1.1.3           R6_2.5.1           
[57] lubridate_1.9.0     knitr_1.41          fastmap_1.1.0       future.apply_1.10.0 utf8_1.2.2          parallel_4.2.2      Rcpp_1.0.9         
[64] vctrs_0.5.1         rpart_4.1.19        tidyselect_1.2.0    xfun_0.35  
