---
title: "Preprocess your data with recipes"
author: "Mike Kaminski"
date: "2023-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE,
                      message = FALSE, dpi = 180,
                      fig.width = 10, fig.height = 5)
```

# Intro
Previously we learned how to specify and train models using different engines from parsnip.  This workbook will review the recipes package, which is designed to help with preprocessing the data before training.
 * creating dummy variables
 * transforming/scaling data
 * transforming groups of variables
 * extracting key features from variables
 
## Flights Data
The goal is to predict if a flight is going to be more than 30 mins late
```{r}
library(tidymodels) # broom, dials, dplyr, ggplot2, infer, parsnip, purrr, recipes, rsample, tibble, tune, workflows, yardstick

# Helper packages
library(nycflights13)    # for flight data - variable named "flights", "weather",
library(skimr)           # for variable summaries
```
 
```{r}
set.seed(123)

flight_data <- 
  flights %>% 
  mutate(
    # Convert the arrival delay to a factor
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    
    # We will use the date (not date-time) in the recipe below
    date = lubridate::as_date(time_hour)
  ) %>% 
  
  # Include the weather data
  inner_join(weather, by = c("origin", "time_hour")) %>% 
  
  # Only retain the specific columns we will use
  select(dep_time, flight, origin, dest, air_time, distance, 
         carrier, date, arr_delay, time_hour) %>% 
  
  # Exclude missing data
  na.omit() %>% 
  
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
  mutate_if(is.character, as.factor)
```

```{r}
flight_data %>% 
  count(arr_delay) %>% 
  mutate(prop = n/sum(n))

```
About 16% of flights are delayed by 30mins or more

A quick glimpse of the data shows us the variables.  We want to predict if a flight is more than 30mins late, so the arr_delay variable is the variable we to predict using logistic regression.

There are two variables that we don't want to use as predictors, but want to retain - flight and time_hour.

There are 104 different destinations and 16 different carries.  We need to convert these to dummy variables.  However, some of these values are infrequent, which could significantly impact the analysis.  This will need to be addressed before modeling.
```{r}
glimpse(flight_data)

flight_data %>%
  skimr::skim(dest, carrier)
```

# Splitting the data
A training and testing data set are needed in order to test the model.  It might be useful to also create a validation set, but that could potentially be addressed later.

rsample has a function called initial_split that can be used to split the data. training and testing functions are used to extract the appropriate data
```{r}
set.seed(234)
data_split <- initial_split(flight_data, prop = 3/4)

train_data <- training(data_split)
test_data <- testing(data_split)

```

# Creating Recipes and Roles
We'll start by creating a simple logistic regression model.  Before training, we can use the recipe to create new predictors and do some preprocessing of the data.

The update_role function allows us to retain some of the variables, but not include them in the model.  We can call these variables "ID".  The data is retained in case we need to go back and investigate some of the variables.  The summary function provides the assigned roles for each of the variables.
```{r}
flights_rec <-
  recipe(arr_delay ~ ., data = train_data) %>% 
  update_role(flight, time_hour, new_role = "ID") 

summary(flights_rec)

```

# Creating Features
The date variable in the flights data could potentially provide us with some value - day of the week, the month, if the date was a holiday.  It's likely that airport behavior varies by day of the week - weekends, by month - flights to warmer regions, or by holidays - holiday travel.  The step_date and step_holiday function allow us to extract those features from these variables.  Step_dummy is used to create dummy variables for the nominal predictors.

There are a ton of different step_functions - found here: https://recipes.tidymodels.org/reference/index.html
  * imputing values (impute_mean, impute_knn, impute_linear)
  * transformations (log, relu, logit)
  * dummy variables (count, dummy, holiday,regex)
  * interactions
  * normalization (mutate_at, pca)
  * row operations (arrange, filter, lag)

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>% 
  update_role(flight, time_hour, new_role = "ID") %>% 
  step_date(date, features = c("dow", "month")) %>%               
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"), 
               keep_original_cols = FALSE) %>% 
  step_dummy(all_nominal_predictors())
```

Given that carrier and dest have some infrequent occurrences, it would be helpful to remove those values that aren't in both data sets.  In this instance, all carriers and dests are present in both data sets.
  * LEX only has one value, so we can remove that since the single value can't be in both the train and test data set.
  * this can be done using the step_zv() function - which removes columns from the data with only a single value.  The dummy values were created previously, so step_zv() looks at all the new binary columns.
```{r}
flight_data %>%
  group_by(dest) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  arrange(count) %>%
  summary()
  #ggplot(aes(count)) + geom_boxplot()

flight_data %>%
  group_by(carrier) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  arrange(count) %>%
  summary()
  #ggplot(aes(count)) + geom_boxplot()
```

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>% 
  update_role(flight, time_hour, new_role = "ID") %>% 
  step_date(date, features = c("dow", "month")) %>%               
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"), 
               keep_original_cols = FALSE) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_zv()

```

# Fit a model with a recipe
For this model, we'll use glm in the set_engine.  The recipe will be used across several steps as the model is trained and tested.
  * Process the recipe using the training data: converting to dummy variables and removing zero variance variables.
  * Apply the recipe to the training data: a final predictor set is created based on the training data
  * Apply the recipe to the testing data: the flights recipe has all that we need to apply to both the train and test datasets.  No other computations need to be made, just need to update the code.
```{r}
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")
```

To do this, a workflow needs to be created - which pairs the model and the recipe together.  It's a straightforward approach given that differnet recipes are often needed for different models.  When a model and recipe are bundled, it makes it easier to train and test workflows.
```{r}
flights_wflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flights_rec)
```

A single function can be used to prepare the recipe and train the model.
```{r}
flights_fit <- 
  flights_wflow %>% 
  fit(data = train_data)
```

The extract_fit_parsnip function extracts the model data, while the tidy() function makes the results a tibble
```{r}
flights_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

```

# Making Predictions
We've built the model, preprocessed the data using recipes, bundled the model and recipe with a workflow, and trained the workflow.  Now it's time to make predictions on the data.
```{r}
# two different ways to show the data
predict(flights_fit, test_data)
predict(flights_fit, test_data, type = "prob")
```
A third way to show the data
```{r}
flights_aug <- 
  augment(flights_fit, test_data)

flights_aug %>%
  select(arr_delay, time_hour, flight, .pred_class, .pred_on_time)

```

We have predictions, but now we need to evaluate the predictions.  The yardstick pacakage includes roc_curve() and roc_auc().  Auto_plot() can help with plotting
```{r}
flights_aug %>% 
  roc_curve(truth = arr_delay, .pred_late) %>% 
  autoplot()

flights_aug %>% 
  roc_auc(truth = arr_delay, .pred_late)
```

